---
title: "Feiyaw__OriginalHomeworkCode_04"
author: "Feiya Wang"
date: "10/23/2019"
output: html_document
---
***

## Homework 4

***
Introduction: This is the code for [Homework 04](https://fuzzyatelin.github.io/bioanth-stats/homework-04.html), on week 7 of AN597

# Question 1
>Write a simple R function, Z.prop.test(), that can perform one- or two-sample Z-tests for proportion data, using the following guidelines.

1.Your function should take the following arguments: p1 and n1 (no default) representing the estimated proportion and sample size (i.e., based on your sample data); p2 and n2 (both defaulting to NULL) that contain a second sample’s proportion and sample size data in the event of a two-sample test; p0 (no default) as the expected value for the population proportion; and alternative (default “two.sided”) and conf.level (default 0.95), to be used in the same way as in the function t.test().

2.When conducting a two-sample test, it should be p1 that is tested as being smaller or larger than p2 when alternative=“less” or alternative=“greater”, the same as in the use of x and y in the function t.test().

3.The function should perform a one-sample Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

4.The function should contain a check for the rules of thumb we have talked about (n∗p>5 and n∗(1−p)>5) to ensure the validity of assuming the normal distribution in both the one- and two-sample settings. If this is violated, the function should still complete but it should also print an appropriate warning message.

5.The function should return a list containing the members Z (the test statistic), P (the appropriate p value), and CI (the two-sided CI with respect to “conf.level” around p1 in the case of a one-sample test and around p2-p1 in the case of a two-sample test). For all test alternatives (“two.sided”, “greater”, “less”), calculate symmetric CIs based on quantiles of the normal distribution rather than worrying about calculating single-limit confidence bounds.



```{r}
z.prop.test <-function(p1,n1,p2=NULL,n2=NULL,p0,alternative=c("two.sided", "less", "greater"),conf.level=0.95){
  #seeting function for allthe required default and no default values
  m=mean(p1)
  #mean of population one
  ste=sqrt(p0(1−p0)/n1)                  
  #standard error of population one.
  s=sd(p1)
  #standard deviation of p1
  a=1-conf.level
  # alpha value
  
  if(is.NULL(p2) | is.NULL(n2)){  # if p2 and n2 are null, we do the one-propotion z test.
      if(n1∗p0>5 & n1∗(1−p0)>5){    #check for rule of thumbs
        z=(m-p0)/ste                       
        # z-score calculation
      
        if(alternative=="less"){      # contain if(less), else if(greater), else if(two sided), else(error)
         p <- pnorm(z)
         # p value calculation
         
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
          test=FALSE
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
      }
    
        else if(alternative=="greater"){
          p <- 1-pnorm(z)
          # p value calculation
        
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
          test=FALSE
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
      }
      
        else if(alternative=="two.sided"){ 
          p <- 2*pnorm(z)
          #p value calculation
        
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
          test=FALSE
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
    }
    
        else { #error
       warning("it is a one-proptional z test")
      # error message
    }
      
      }
      else{     # cannot pass rule of thumbs
      warning("cannot assume this normal approximation since it didn't pass rule of thumb check")
      }
    }
    
  else if(!(is.null(p2)&is.null(n2))){  "if we have p2,n2 value, we should do two proption z test"
      m1=mean(p1)
      #mean of population one
      m2=mean(p2)
      #mean of population two
      p0=(sum(p1)+sum(p2))/(n1+n2)
      # define the pooled propotion
      ste=sqrt(p0*(1−p0)*(1/n1+1/n2))
      # standard error of p1+p2
      if(n1∗p0>5 & n1∗(1−p0)>5 & n2∗p0>5& n2∗(1-p0)>5){            #check for rule of thumbs
        z=(m1-m2)/ste
        # z-score for two propotion z test
        m= m1-m2
        #restate m. 
        if(alternative=="less"){   # contain if(less), else if(greater), else if(two sided), else(error)
          p <- pnorm(z)
          # p value calculation
          
          
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
          test=FALSE
          
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
      }
    
        else if(alternative=="greater"){
            p <- 1-pnorm(z)
            # p value calculation
            
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
         
              test=FALSE
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
      }
      
        else if(alternative=="two.sided"){ 
          p <- 2*pnorm(z)
          # p value calculation
          
         lower <- m - qnorm(1 - a/2) * ste  
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         upper <- m + qnorm(1 - a/2) * ste 
         # (1-alpha)/2 each in the upper and lower tails of the distribution
          
         ci <- c(lower, upper)
         # confidence interval
          
         test=TRUE
         #result of test as a boolean veriable.
           
           r=c(z,p,ci,test)
           # variable for return
        
           if(p< a){        # check if p value is less than alpha value.if(yes), else (no) 
          
         
             return(r)
         }
           else{
          test=FALSE
          return(r)
          # cannot reject hypothesis because p value > alpha
        }
    }
    
        else { #error
       warning("it is a one-proptional z test")
      # error message
    }
      
      }
      else{    # cannot pass the check for rule of thumbs
        warning("cannot assume this normal approximation since it didn't pass rule of thumb check")
      }
    }
  
  else {
        warning("cannot assume this normal approximation since it didn't pass rule of thumb check")
    }
  }

```

Here is the test of the z.prop.function we write.
```{r}

```



# Question 2

>The dataset from Kamilar and Cooper has in it a large number of variables related to life history and body size. For this exercise, the end aim is to fit a simple linear regression model to predict longevity (MaxLongevity_m) measured in months from species’ brain size (Brain_Size_Species_Mean) measured in grams. Do the following for both longevity~brain size and log(longevity)~log(brain size):

```{r}
library (curl)
#install package curl before this step to load a file from server.
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall19/KamilarAndCooperData.csv")
#load the file in varibale f
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
# read file in d as a dataframe.
```


>1.Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r}
p1 <- ggplot(data=d, aes(x=MaxLongevity_m,y=Brain_Size_Species_Mean))
# build a ggplot object
p1 <- p1 + geom_point() 
# make a scatterplot
p1 <- p1 + geom_smooth(method="lm", fullrange=TRUE)
# add a regression line 
m<-lm(data=d,MaxLongevity_m ~ Brain_Size_Species_Mean)
# make the model of two variable
t <- unlist(m$coefficients)
# unlist m to get coeffient
beta0<-round(t[1],digits = 3)
#round beta 1 to 3 decimal place
beta1<-round(t[2],digits = 3)
#round beta 0 to 3 decimal place
soe1<-paste("y=",as.character(beta1),"x+",as.character(beta0))

p1<-p1+annotate("text", label = soe, x = 600, y = 500, size = 5, colour = "black")
p1

r1<-log(d$MaxLongevity_m)
r2<-log(d$Brain_Size_Species_Mean)

p2 <- ggplot(data=d, aes(x=r1,y=r2))
# build a ggplot object
p2 <- p2 + geom_point() 
# make a scatterplot
p2 <- p2 + geom_smooth(method="lm", fullrange=TRUE)
# add a regression line 
m<-lm(data=d,log(MaxLongevity_m) ~ log(Brain_Size_Species_Mean))
# make the model of two variable
t <- unlist(m$coefficients)
# unlist m to get coeffient
beta0<-round(t[1],digits = 3)
#round beta 1 to 3 decimal place
beta1<-round(t[2],digits = 3)
#round beta 0 to 3 decimal place
soe2<-paste("y=",as.character(beta1),"x+",as.character(beta0))

p2<-p2+annotate("text", label = soe, x = 600, y = 500, size = 5, colour = "black")
p2

```

>2.Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

```{r}

```

>3.Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.

```{r}

```

>4.Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
Looking at your two models, which do you think is better? Why?

```{r}

```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

